---
title: "lab1"
output: html_document
date: "2026-09-17"
---

# Overview

1. Create a df of the Airquality dataset. Display names of the columns. Check data types of the attributes. Run a summary statistics of the dataset.

2. Check the missing values in the dataset and display NAs per column

3. Day and month are seen as integers. Make them factors. Check the data types of the attributes

4. Calculate the mean of the non-missing values in the Ozone column. Replace the missing values with this mean. Do the same for Solar.R

5. Use a histogram to assess normality of the Ozone variable. Does it appear to be normally distributed?

6. Create a boxplot which shows the distribution of Ozone in each month. Use different colours for each month

7. Create a scatter plot matrix of the numeric variables within the Airquality dataset

# Solution

## Load the libraries
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(janitor)
```

## Answer the questions

1. Create a df of the Airquality dataset. Display names of the columns. Check data types of the attributes. Run a summary statistics of the dataset.
```{r}
head(airquality) # already available in R

df <- airquality |> clean_names() # create df

colnames(df) # display names

str(df) # check data types of attributes with str()

summary(df) # see summary stats of the df with summary()
```

2. Check the missing values in the dataset and display NAs per column
```{r}
df |> summarize(
  across(everything(), ~ sum(is.na(.)))
)
```

3. Day and month are seen as integers. Make them factors. Check the data types of the attributes
```{r}
# convert int to factor
df$month <- as.factor(df$month)
df$day <- as.factor(df$day)

str(df)
```

4. Calculate the mean of the non-missing values in the Ozone column. Replace the missing values with this mean. Do the same for Solar.R
```{r}
oz_mean <- mean(df$ozone, na.rm = TRUE)

# where oz is.na(), replace with oz_mean

solar_mean <- mean(df$solar_r, na.rm = TRUE)
```

5. Use a histogram to assess normality of the Ozone variable. Does it appear to be normally distributed?
```{r}
df |>
  select(ozone) |>
  ggplot(aes(x=ozone)) +
    geom_histogram()

```

6. Create a boxplot which shows the distribution of Ozone in each month. Use different colours for each month
```{r}
df |>
  select(ozone, month) |>
  ggplot(aes(x=ozone, y=month, fill=month)) +
    geom_boxplot()
```

7. Create a scatter plot matrix of the numeric variables within the Airquality dataset
```{r}
df |>
  select(ozone, solar_r, wind, temp) |>
  ggplot() +
    geom_point()
```


